\Large UC President's Postdoctoral Fellowship Research Statement: \\ \textbf{How Unique is the Solar System? Planetary Luminosities as Fingerprints for the Origins and Properties of Exoplanets}

### 2016-11-06 Postdoc Applications/uc_background_statement.tex
upd

I was born in Romania a couple of years before the 1989 revolution. As a woman growing up in a country that was not ready to promote values such as diversity, inclusion, tolerance, and equality between genders, I was not expected to pursue a scientific career. My family supported my choice, but society and the educational system did not. I moved to the United States to attend college with the expectation that the values and ethics that are often disregarded in my home country would be universally appreciated here. Unfortunately, I found out that this is not always the case, especially regarding gender equality. As I gradually advanced in my career in astrophysics, this disparity became more and more evident. As did the fact that girls and young women are often discouraged from considering scientific career paths, either due to the scarcity of opportunities and resources, or their own lack of self confidence, galvanized by little societal support or very few role models. I am therefore highly dedicated to mentoring, encouraging and empowering women, a group often deeply underrepresented in scientific fields.

Throughout my graduate studies, I have often been a mentor, in various capacities, to younger women already pursuing a scientific career, or just interested in and excited about the possibility. I was part of the Women In Science, Engineering, Technology and Mathematics (WISTEM) organization, a group that aims to foster a sense of community for women studying science and engineering at Harvard. The program pairs graduate and undergraduate students in a mentor-mentee relationship, with the aim of creating a network for women scientists. I served as a mentor for three years to a Harvard undergraduate. Through monthly coffees or lunches, we discussed graduate school opportunities, work/life balance, some of the challenges that she faced both as a woman and a minority, and the general prospect of a career in science. She had successful summer internships at Harvard Medical School and Harvard School of Public Health, and is now in her final year as a biostatistics major. In a similar capacity, I have been a mentor for a summer Research Experiences for Undergraduates (REU) student at the Center for Astrophysics (CfA). As part of Prof. Karin \"Oberg's group, I co-mentored two of her female undergraduate students for their senior thesis projects, one of whom received the award for best senior thesis, and the other who is now a graduate student at the CfA.

>>>>

I was born in Romania a few years before the 1989 revolution. As a woman growing up in a country that was not ready to promote values such as diversity, inclusion, tolerance, and equality between genders, I was not expected to pursue a scientific career. My family supported my choice, but society and the educational system did not. I moved to the United States to study Physics at MIT with the expectation that the values and ethics that are often disregarded in my home country would be universally appreciated here. Unfortunately, I found out that this is not always the case, especially regarding gender equality. As I gradually advanced in my career in astrophysics, this disparity became more and more evident. As did the fact that girls and young women are often discouraged from considering scientific career paths, either due to the scarcity of opportunities and resources, or their own lack of self confidence, galvanized by little societal support or very few role models. I am therefore highly dedicated to mentoring, encouraging and empowering women, a group often deeply underrepresented in scientific fields.

Throughout my graduate studies at Harvard University, I have often been a mentor, in various capacities, to younger women already pursuing a scientific career, or just interested in and excited about the possibility. I was part of the Women In Science, Engineering, Technology and Mathematics (WISTEM) organization, a group that aims to foster a sense of community for women studying science and engineering at Harvard. The program pairs graduate and undergraduate students in a mentor-mentee relationship, with the aim of creating a network for women scientists. I served as a mentor for three years to a Harvard undergraduate. Through monthly coffees or lunches, we discussed graduate school opportunities, work/life balance, some of the challenges that she faced both as a woman and a minority, and the general prospect of a career in science. She had successful summer internships at Harvard Medical School and Harvard School of Public Health, and is now in her final year as a biostatistics major. In a similar capacity, I have been a mentor for a summer Research Experiences for Undergraduates (REU) student at the Center for Astrophysics (CfA). As part of Prof. Karin \"Oberg's group, I co-mentored two of her female undergraduate students for their senior thesis projects, one of whom received the award for best senior thesis, and the other who is now a graduate student at the CfA.

### 2016-11-06 Postdoc Applications/uc_research_statement.tex
upd

Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered [1]. Their diversity in terms of mass, radius, location and composition [2] provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and may sustain life. The advent of the Kepler Space Telescope has resulted in an exponential increase in the number of exoplanets detected via the transit method. As shown in Figure 1, a large fraction of the planets detected by Kepler are the so-called ``Super Earths'' or ``Mini Neptunes'', planets with masses between 1 and 10 times that of the Earth, and which are likely to contain some of their mass in a gaseous envelope. The majority of these detected Super Earths are very close to their host stars, with smaller orbital periods than the Earth. These features imply that Super Earths do not have Solar System analogs. At the same time, they are ubiquitously found and predicted to be the most common types of exoplanets [3], which makes understanding their ages, formation and properties particularly relevant. Observational techniques such as gravitational microlensing and direct imaging have extended the orbital range at which planets have been detected from just outside the Earth's orbit to tens and even hundreds of astronomical units. As demonstrated in Figure 2, however, most of these planets are massive; wide-separation Super Earths or Neptune analogs have yet to be detected. Are they indeed rare at these large distances, or just not bright or young enough to be detected? The Gemini Planet Imager (GPI) exoplanet survey will be key in improving our knowledge on this matter. GPI is a direct imaging survey that aims to observe very young, bright planets. As planets become less bright with increasing age, GPI might thus be able to detect lower mass planets that are just at the end stages of their formation. A theoretical framework to interpret these observations is therefore crucial. In particular, a planet's luminosity is an essential feature because it can provide constraints on the planet's mass, age and formation mechanism. \textbf{I propose to develop a holistic framework to self consistently track a planet's luminosity from its formation and through its evolution after the dispersal of the protoplanetary disk. Such a model provides essential context for characterizing young planets that are and will be discovered via direct imaging surveys.}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{exo_solar}
%\vspace{-0.5in}
\caption{Planetary mass vs. orbital separation of the detected exoplanet population, via transit (red) and radial velocity (blue). Venus, Mars and Earth are displayed in the lower right corner. No Solar System Super Earth analogs have been yet detected.}
\label{fig:chemical}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{exo_solar_all}
%\vspace{-0.5in}
\caption{Same as above, but planets detected via gravitational microlensing (green) and direct imaging (yellow) are added. Besides the terrestrial planets, Jupiter, Saturn, Uranus and Neptune are also shown. There is a scarcity in Super Earth detections at wide separateions.}
\label{fig:chemical}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{opacity_effect.pdf}
%\vspace{-0.5in}
\caption{Example of a planet's luminosity and time evolution with atmospheric mass, during the giant planet formation stage. From [6].}  %The actual magnitude of dust opacities is difficult to robustly predict.}
\label{fig:LtvsMopacity}
\end{figure}

@@

%\section{Model Planet Populations} 
\textbf{3. Luminosity evolution and connection to observations.} A planet's luminosity evolution has been analyzed before [11], but this was only done for gas giants. Moreover, these models consider arbitrary initial conditions, as they are only interested in studying the luminosities of mature, several billion years old exoplanets. \textbf{The combined results from parts 1 and 2 will represent the first time that a planet's luminosity evolution is tracked self consistently from its birth through after the disk dispersal.} These results will be crucial to understanding the properties of directly imaged young, bright planets. Furthermore, they may be able to help us distinguish between different formation scenarios Ð core accretion versus gravitational instability.

@@

\textbf{UCLA is the ideal place for me to continue my postdoctoral tenure as a President's Fellow.} I have tremendously enjoyed the time that I have spent here, and greatly benefitted from the diversity and academic excellence of the EPSS department. In particular, working with Prof. Hilke Schlichting is a privilege, and her expertise in planet formation and dynamics will be instrumental in achieving my research goals as a postdoctoral fellow and afterwards. Collaborating with Prof. David Jewitt, a leader in Solar System astrophysics, will be very useful for analyzing my findings in the context of the Solar System. As my research interests are at the intersection of astrophysics and planetary science, I would welcome collaborations with members of the Astronomy department faculty, such as Prof. Smadar Naoz. UCLA is also home to the Institute for Planets and Exoplanets (iPLEX), an interdisciplinary forum that ties together people with different scientific backgrounds. Outside research, I plan to be involved in the UCLA Social Postdoc Society. Through this, I will help foster connections between postdoctoral scholars from various departments across the campus, with the goal of sharing our diverse scientific knowledge and experience.
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}

>>>>

Within the last two decades, more than one thousand extrasolar planets (exoplanets) have been discovered [1]. Their diversity in terms of mass, radius, location and composition [2] provides an exciting field of research, with the eventual goal of finding planets that are similar to our own Earth and that may sustain life. The advent of the Kepler Space Telescope has resulted in an exponential increase in the number of exoplanets detected via the transit method. As shown in Figure 1, a large fraction of the planets detected by Kepler are the so-called ``Super Earths'' or ``Mini Neptunes'', planets with masses between 1 and 10 times that of the Earth, and which are likely to contain some of their mass in a gaseous envelope. The majority of these detected Super Earths are very close to their host stars, with smaller orbital periods than the Earth. These features imply that Super Earths are not Solar System analogs. At the same time, they are ubiquitously found and predicted to be the most common types of exoplanets known to date [3], which makes understanding their ages, formation and properties particularly relevant. Observational techniques such as gravitational microlensing and direct imaging have extended the orbital range at which planets have been detected from just outside the Earth's orbit to tens and even hundreds of astronomical units. As demonstrated in Figure 2, however, most of these planets have Jupiter-like masses; true Uranus or Neptune analogs have yet to be detected. Are they indeed rare at these large distances, or just not bright or young enough to be detected? The Gemini Planet Imager (GPI) exoplanet survey will be key in improving our knowledge on this matter. GPI is a direct imaging survey that aims to observe very young, bright planets. As planets become less bright with increasing age, GPI might thus be able to detect lower mass planets that are just at the end stages of their formation. A theoretical framework to interpret these observations is therefore crucial. In particular, a planet's luminosity is an essential feature because it can provide constraints on the planet's mass, age and formation mechanism. \textbf{I propose to develop a holistic framework to self consistently track a planet's luminosity from its formation and through its evolution after the dispersal of the protoplanetary disk. Such a model provides essential context for characterizing young planets that are and will be discovered via direct imaging surveys.}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{exo_solar}
%\vspace{-0.5in}
\caption{Planetary mass vs. orbital separation of the detected exoplanet population, via transit (red) and radial velocity (blue). Venus, Mars and Earth are displayed in the lower right corner. The detected Super Earths have no Solar System analogs.}
\label{fig:chemical}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.8\textwidth]{exo_solar_all}
%\vspace{-0.5in}
\caption{Same as above, but planets detected via gravitational microlensing (green) and direct imaging (yellow) are added. Besides the terrestrial planets, Jupiter, Saturn, Uranus and Neptune are also shown. There is currently a scarcity of exoplanets with Earth to Saturn Masses outside $\sim$5 AU. New direct imaging surveys are expected to fill in this phase space.}
\label{fig:chemical}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.5\textwidth]{opacity_effect.pdf}
%\vspace{-0.5in}
\caption{Example of a planet's luminosity and time evolution with atmospheric mass, during the giant planet formation stage. The different colored curves correspond to different assumptions about the atmospheric opacity. From [6].}  %The actual magnitude of dust opacities is difficult to robustly predict.}
\label{fig:LtvsMopacity}
\end{figure}

@@

%\section{Model Planet Populations} 
\textbf{3. Luminosity evolution and connection to observations.} A planet's luminosity evolution has been analyzed before [11], but this was only done for gas giants. Moreover, these models consider arbitrary initial conditions, as they are only interested in studying the luminosities of mature, several billion years old exoplanets. \textbf{The combined results from parts 1 and 2 will represent the first time that a planet's luminosity evolution is tracked self consistently from its birth through after the disk dispersal.} These results will be crucial for determining the masses of directly imaged exoplanets. They will be essential for correctly extrapolating from the observations of young, and hence bright systems, to the underlying population of exoplanets that is dominated by older, and hence significantly less luminous planets that will remain undetectable. Furthermore, these results may be able to help us distinguish between different formation scenarios --- core accretion versus gravitational instability.

@@

\textbf{UCLA is the ideal place for me to continue my postdoctoral tenure as a President's Fellow.} I have tremendously enjoyed the time that I have spent here, and greatly benefitted from the diversity and academic excellence of the EPSS department. In particular, working with Prof. Hilke Schlichting is a privilege, and her expertise in planet formation and dynamics will be instrumental in achieving my research goals as a postdoctoral fellow and afterwards. Collaborating with Prof. David Jewitt, a leader in Solar System astrophysics, will be very useful for analyzing my findings in the context of the Solar System. As my research interests are at the intersection of astrophysics and planetary science, I would welcome collaborations with members of the Astronomy department faculty, such as Prof. Smadar Naoz. UCLA is also home to the Institute for Planets and Exoplanets (iPLEX), an interdisciplinary forum that ties together people with different scientific backgrounds. %Outside research, I plan to be involved in the UCLA Social Postdoc Society. Through this, I will help foster connections between postdoctoral scholars from various departments across the campus, with the goal of sharing our diverse scientific knowledge and experience.
%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.7\textwidth]{CO_abundances}
%%\vspace{-0.5in}
%\caption{TBD}
%\label{fig:CO}
%\end{figure}
